---
title: "Prediction Assignment Writeup"
output: html_document
---
# Peer-graded Assignment : Prediction Assignment Writeup #
# Harnoor Rangi#


## Project Background ##
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data sources for the assignment ##

The training data is downloaded from the following link and saved on to the local machine : 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data is downloaded fromthe following link and saved on to the local machine:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Loading data ##

Data files from the above link were downloaded to the local computer and loaded into R using the following code:
```{r}
train_data <-read.csv("pml-training.csv")
test_data <-read.csv("pml-testing.csv")
```

## Loading required libraries##
```{r}

library(caret)
library(rattle)
library(lattice)
library(rpart)
library(rpart.plot)
library(RGtk2)
library(RColorBrewer)
library(randomForest)
library(e1071)

```

## Data Cleaning ##
Removing the rows of data containing "NA" and the variables which are useless for prediction from both training and test data sets
```{r}
set.seed(12345)
unwanted <- names(test_data[,colSums(is.na(test_data)) == 0])[8:59]

train_data <- train_data[,c(unwanted,"classe")]
test_data <- test_data[,c(unwanted,"problem_id")]

dim(train_data)
dim(test_data)
```

## Spliting the data as per requirement##
From the training data set creating partitions to create two new data sets 60 % training and 40% testing
```{r}
part <- createDataPartition(train_data$classe, p = 0.6, list=FALSE)
new_train <- train_data[part,]
new_test <- train_data[-part,]

dim(new_train)
dim(new_test)
```

## Building the model ##

###Creating model using Random Forest first###

```{r}
modelfit1 <- randomForest(classe~., data=new_train, method='class')
prediction1 <- predict(modelfit1,new_test,type='class')
qplot(roll_belt, magnet_dumbbell_y, colour=classe, data=new_train)

x<- confusionMatrix(prediction1,new_test$classe)
x$table
x$overall[1]
```

Hence, we get 99% accuracy with Random forest method. 

###Creating Decision tree model### 
```{r}
modelfit2<- rpart(classe ~ .,data=new_train, method='class')
fancyRpartPlot(modelfit2)

prediction2 = predict(modelfit2,new_test,type='class')
y<-confusionMatrix(prediction2,new_test$classe)
y$table
y$overall[1]
```

## Cleary, Random Forest has more accuracy of 99%##

## Using Random forest to predict the test data set ##
```{r}
predictiontest <- predict(modelfit1,test_data,type='class')
predictiontest
```

